---
title: "Statistical Inference Project"
author:  Mark Sucato
output: 
  pdf_document:
    keep_md: true 
---
```{r include=FALSE}
knitr::opts_chunk$set(cache=TRUE, warning=FALSE, message=FALSE, error=FALSE,
	results="asis")
```
<center>
# Statistical Inference Project  
### by Mark Sucato
### *** Complete R code provided in Appendix 1 ***
</center>

```{r, echo=FALSE}
library(knitr)
library(car)
library(datasets)
library(tidyverse)
library(kableExtra)
library(cowplot)
set.seed(1234)
```

### Overview

blah blah blah

<center>
## Part One:  Simulation of Exponential Distributions
</center>

### Simulation specifics and descriptive statistics discussion

Simulation of 1,000 exponential distribution samples (n=40 per sample, lambda =0.2) is 
achieved through a single-line 'for' loop that creates a vector *results* 
containing the mean of each n=40 sample.

```{r}
results <- NULL
for(i in 1:1000) results[i] <- mean(rexp(40, 0.2))
```

Actual and theoretical means and variances are shown below in *Table 1*. As expected, the mean of the sample distribution means closely aligns to the underlying 
exponential distribution mean.  This validates the concept of the sample mean as a 
consistent estimator of the population mean. The sample distribution variance, however, does
not closely align to the the theoretical sample population variance.  This variance 
is a function of the underlying exponential distribution variance and the size of the 
n=40 samples.  With more simulation repetitions, this sample distribution variance would more closely 
approach the theoretical value.  

```{r, echo=FALSE}
mu <- 1/0.2         
xBar <- round(mean(results), digits = 3)
varExp <- 1/(0.2)^2     
varBar <- round(varExp/40, digits = 3)       
varActual<- round(var(results), digits = 3)  
Value <- c(mu, xBar, varBar, varActual)
Stats <- c("Population Mean", "Sample Mean", "Theoretical Sample Variance", "Actual Sample Variance")
tableData <- cbind(Stats, Value)
wideTable <- pivot_wider(as_tibble(tableData), names_from=Stats, values_from = Value)
kable(wideTable, align = "cccc", caption = 'Table 1: Comparison of Means and Variances') %>%
	kable_styling(position = "center") %>%
	kable_classic(full_width = F)
```

### Simulation distribution normality

A histogram of the simulation distribution, with normal distribution superimposed,
 is shown below in *Figure 1A*. A QQ plot of the simulation distribution is shown
in *Figure 1B*.

```{r, echo=FALSE}
h <- hist(results, breaks=20, main = "Figure 1: Distribution of 1000 Simulated Exponential Sample Means
	 (n=40)", xlab = "Means of Exponential Samples (n=40)", col="firebrick")
xfit <- seq(min(results), max(results), length=length(results))
yfit <- dnorm(xfit, mean=mean(results), sd=sd(results))
yfit <- yfit*diff(h$mids[1:2])*length(results)
p1 <- lines(xfit, yfit, col="blue", lwd=2)
p2 <- qqPlot(results, ylab = "Simulated Distribution", xlab = "Normal Distribution",
	main = "QQ Plot for 1000 Simulated Exponential Sample Means",
	id = FALSE)
plot_grid(p1, p2, labels = c('1A', '1B'), label_size = 12)

```

 





<center>
## Part Two:  Exploration of ToothGrowth Dataset  
</center>

